Full Analysis Script

### 2024-09-06 code/analysis-code/fullanalysis-v1.qmd
Final Revisions for Submission
Cleaned up all files

```{r, message=FALSE,  warning=FALSE}
#Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear.rds")
#load data
ML_linear <- readRDS(data_location)
str(ML_linear)
```

@@

First, I will create train and test data (75/25) to be able to evaluate the performance of our models. I will also set the seed to ensure reproducibility. For each model, I perform the following steps: 1) run the model on the train data; 2) perform cross-validation on the train data; 3) evaluate the model on the test data. The model outputs included in the final manuscript are on all the data.

@@

whereas $\alpha$ is the average birth weight, $\beta_1$ is the group deviation from the average birth weight, `BW` is birth weight in grams, `MA` is maternal ancestry (Euro-descent, Afro-descent), and $\epsilon$ is the error term.

@@

```{r}
# Model 1: Final model for paper using all data
# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable. Cannot do this if I am merging tables

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred <- bind_cols(ML_fit1_pred, ML_linear %>% select(Weightgrams))
ML_fit1_pred

@@

#Print table
table_model1

# Save regression output into table
saveRDS(table_model1, file = here::here("results", "tables", "table_model1_final.rds"))
```
## Model 1 Performance

First we will evaluate the model on our training data.

```{r}
# Training model 1
# Model using recipes and workflows
# Define recipe for model
ML_recipe1_train <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred_train <- bind_cols(ML_fit1_pred_train, train_data %>% select(Weightgrams))
ML_fit1_pred_train

@@

# Save regression output into table
saveRDS(table_model1_train, file = here::here("results", "tables", "table_model1_train.rds"))
```
Our results here are not that different from the model run on all the data. Our adjusted R-squared is 0.01, the same as the full model. The RMSE is 488, which is slightly lower than the full model (494), which shows better performance.

Now let's perform cross-validation on the training data.

@@

@@

# Save regression output into table
saveRDS(ML_cv1_metrics, file = here::here("results", "tables", "ML_cv1_metrics.rds"))
```
Performing cross-validation results in an RMSE that is the same as our training model (488), lower than our full model. The r-squared (0.017) is also higher than our training and full models (0.01), demonstrating better model performance.

Finally, let's run the model on our test data.

```{r, message=FALSE,  warning=FALSE}
# Testing model 1
# Model using recipes and workflows
# Define recipe for model
ML_recipe1_test <- recipe(Weightgrams ~ ModifiedColor, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit1_pred_test <- bind_cols(ML_fit1_pred_test, test_data %>% select(Weightgrams))
ML_fit1_pred_test

@@

@@

When running the model on our testing dataset, our results remain significant at the 0.05 level. The RMSE is 508 and the r-squared is 0.008. This shows a slightly worse fit than our training results.

@@

whereas $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `BW` is birth weight in grams, `MA` is maternal ancestry (Euro-descent, Afro-descent), `Gest` is gestational status (nulliparous or multiparous), `Age` is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

```{r}
# Model 2: Final model for paper using all data
# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) #%>% # Convert all nominal variables to dummy variables. Cannot do this if I am merging tables in gtsummary
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors. Cannot do this if I am merging tables in gtsummary

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred <- bind_cols(ML_fit2_pred, ML_linear %>% select(Weightgrams))
ML_fit2_pred

@@

# Save regression output into table
saveRDS(table_model2, file = here::here("results", "tables", "table_model2_final.rds"))
```
Our RMSE is 490 and our r-squared is 0.026.

@@

Now let's run the model on our training data.

```{r}
# Training model 2
# Using recipes and workflows
# Define recipe for model
ML_recipe2_train <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred_train <- bind_cols(ML_fit2_pred_train, train_data %>% select(Weightgrams))
ML_fit2_pred_train

@@

# Save regression output into table
saveRDS(table_model2_train, file = here::here("results", "tables", "table_model2_train.rds"))
```
Our RMSE is 483 and our r-squared is 0.03, a slightly better performance than the full model.

Now let's run cross-validation.

@@

@@

# Save regression output into table
saveRDS(ML_cv2_metrics, file = here::here("results", "tables", "ML_cv2_metrics.rds"))
```
Here, our RMSE is 483 and our r-squared is 0.026, the same as the training model.

Finally, let's run our model on the test data.

```{r}
# Testing model 2
# Using recipes and workflows
# Define recipe for model
ML_recipe2_test <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit2_pred_test <- bind_cols(ML_fit2_pred_test, test_data %>% select(Weightgrams))
ML_fit2_pred_test

@@

# Save regression output into table
saveRDS(table_model2_test, file = here::here("results", "tables", "table_model2_test.rds"))
```
Running the model on our test data, we see no significant results at any alpha level.

@@

```{r}
# Model 1: Sub-sample simple model (ML_lm1_sub)
# Create a list to store R-squared values
r_squared_values_sub <- list()

@@

# Save regression output into table
saveRDS(ML_lm1_sub_table, file = here::here("results", "tables", "ML_lm1_sub_table.rds"))
```
```{r}
# Model 2: Sub-sample multivariate model (ML_lm2_sub)
# Filter data based on desired conditions
filtered_data <- ML_linear %>%
  filter(Brazilian == "Brazilian")

@@

```{r}
# Combine the four tables using tbl_merge
table3 <- tbl_merge(
  list(table_model1, ML_lm1_sub_table, table_model2, ML_lm2_sub_table),
  tab_spanner = c("**Model 1**", "**Sub-Model 1**", "**Model 2**", "**Sub-Model 2**")) %>%
  modify_table_body(fun = ~.x %>% arrange(variable))
table3

@@

```{r}
# Model 3: Final model for paper using all data
# Using recipes and workflows
# Simple linear model
# Specify recipe
ML_recipe3 <- recipe(Weightgrams ~ Color, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") #%>% # Set reference level for White
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred <- bind_cols(ML_fit3_pred, ML_linear %>% select(Weightgrams))
ML_fit3_pred

@@

# Save regression output into table
saveRDS(table_model3, file = here("results", "tables", "table_model3_final.rds"))
```

Our RMSE is 494 and our r-squared is 0.01.

@@

Let's now evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 3
# Using recipes and workflows
# Specify recipe
ML_recipe3_train <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred_train <- bind_cols(ML_fit3_pred_train, train_data %>% select(Weightgrams))
ML_fit3_pred_train

@@

@@

Our RMSE is 489 and our r-squared is 0.01, slightly better performance than the full model. Our results are not significantly different from the full model.

Now let's run cross-validation.

@@

@@

# Save regression output into table
saveRDS(ML_cv3_metrics, file = here::here("results", "tables", "ML_cv3_metrics.rds"))
```
Using cross-validation, our RMSE is 487 and our r-squared is 0.01, a similar performance to our training model.

Finally, let's run our model on the test data.

```{r, message=FALSE, warning=FALSE}
# Testing model 3
# Using recipes and workflows
# Specify recipe
ML_recipe3_test <- recipe(Weightgrams ~ Color, data = test_data) %>%
  step_relevel(Color, ref_level = "White")

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit3_pred_test <- bind_cols(ML_fit3_pred_test, test_data %>% select(Weightgrams))
ML_fit3_pred_test

@@

@@

Much like in the other testing models, our results are not significant (except the intercept). Our performance is also worse (RMSE of 507 and r-squared of 0.008).

@@

```{r}
# Model 4: Final model for paper using all data
# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit4_pred <- bind_cols(ML_fit4_pred, ML_linear %>% select(Weightgrams))
ML_fit4_pred

@@

# Save regression output into table
saveRDS(table_model4, file = here("results", "tables", "table_model4_final.rds"))
```
Our RMSE is 490 and our r-squared is 0.026.

@@

Let's now evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Model 4 training
# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_train <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Now we will match predicted values with corresponding observed outcome values
ML_fit4_pred_train <- bind_cols(ML_fit4_pred_train, train_data %>% select(Weightgrams))
ML_fit4_pred_train

@@

# Save regression output into table
saveRDS(table_model4_train, file = here("results", "tables", "table_model4_train.rds"))
```
Our RMSE is 483 and our r-squared is 0.033, a slightly better performance than the full model.

Now let's run cross-validation.

@@

@@

# Save regression output into table
saveRDS(ML_cv4_metrics, file = here::here("results", "tables", "ML_cv4_metrics.rds"))
```
Here, our RMSE is 484 and our r-squared is 0.025, a better performance than both the full model and training model.

Finally, let's run our model on the test data.

>>>>

```{r, message=FALSE,  warning=FALSE}
# Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear.rds")

# Load data
ML_linear <- readRDS(data_location)
str(ML_linear)
```

@@

First, I will create train and test data (75/25) to be able to evaluate the performance of the models. I will also set the seed to ensure reproducibility. For each model, I perform the following steps: 1) run the model on the train data; 2) perform cross-validation on the train data; 3) evaluate the model on the test data. The model outputs included in the final manuscript are on all the data.

@@

whereas `BW` is birth weight in grams, $\alpha$ is the average birth weight, $\beta_1$ is the group deviation from the average birth weight, `MA` is maternal ancestry (Euro-descent, Afro-descent), and $\epsilon$ is the error term.

@@

```{r, message=FALSE,  warning=FALSE}
# Model 1: Final model for paper using all data
# Create a list to store R-squared values
r2_1 <- list()
rmse_1 <- list()

# Model using recipes and workflows
# Define recipe for model
ML_recipe1 <- recipe(Weightgrams ~ ModifiedColor, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable. Cannot do this if I am merging tables

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit1_pred <- bind_cols(ML_fit1_pred, ML_linear %>% select(Weightgrams))
ML_fit1_pred

@@

# Print table
table_model1

# Extract RMSE and R2 for in-line text
r2_1$model1 <- metrics_fit1[metrics_fit1$.metric == "rsq", ".estimate"]
rmse_1$model1 <- metrics_fit1[metrics_fit1$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model1, file = here::here("results", "tables", "table_model1_final.rds"))

# Create a gtsummary table directly from the model with CIs
table_model1_CI <-
  tbl_regression(ML_fit1, exponentiate = FALSE, intercept = TRUE) %>% 
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model1_CI
```

The adjusted R-squared is `r round(r2_1$model1, 3)`, and the RMSE is `r round(rmse_1$model1)`.

## Model 1 Performance

First, I will evaluate the model on the training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 1
# Create a list to store R-squared and RMSE values
r2_1_train <- list()
rmse_1_train <- list()

# Model using recipes and workflows
# Define recipe for model
ML_recipe1_train <- recipe(Weightgrams ~ ModifiedColor, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit1_pred_train <- bind_cols(ML_fit1_pred_train, train_data %>% select(Weightgrams))
ML_fit1_pred_train

@@

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_1_train$model1_train <- metrics_fit1_train[metrics_fit1_train$.metric == "rsq", ".estimate"]
rmse_1_train$model1_train <- metrics_fit1_train[metrics_fit1_train$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model1_train, file = here::here("results", "tables", "table_model1_train.rds"))
```

The results are similar to those from the model run on all the data. The adjusted R-squared is `r round(r2_1_train$model1_train, 3)`, the same as the full model. The RMSE is `r round(rmse_1_train$model1_train)`, which is slightly lower than the full model, `r round(rmse_1$model1)`, which shows better performance.

Now, I will perform cross-validation on the training data.

@@

# Create a list to store R-squared and RMSE values
r2_1_cv <- list()
rmse_1_cv <- list()

@@

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_1_cv$model1_cv <- ML_cv1_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_1_cv$model1_cv <- ML_cv1_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv1_metrics, file = here::here("results", "tables", "ML_cv1_metrics.rds"))
```
Performing cross-validation results in an RMSE that is the same as the training model `r round(rmse_1_cv$model1_cv)`, lower than the full model. The r-squared `r round(r2_1_cv$model1_cv, 3)` is also higher than the training `r round(r2_1_train$model1_train, 3)` and full models `r round(r2_1$model1, 3)`, demonstrating better model performance.

Finally, I'll run the model on the test data.

```{r, message=FALSE,  warning=FALSE}
# Testing model 1
# Create a list to store R-squared and RMSE values
r2_1_test <- list()
rmse_1_test <- list()

# Model using recipes and workflows
# Define recipe for model
ML_recipe1_test <- recipe(Weightgrams ~ ModifiedColor, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(ModifiedColor) # Convert ModifiedColor to dummy variable

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit1_pred_test <- bind_cols(ML_fit1_pred_test, test_data %>% select(Weightgrams))
ML_fit1_pred_test

@@

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_1_test$model1_test<- metrics_fit1_test[metrics_fit1_test$.metric == "rsq", ".estimate"]
rmse_1_test$model1_test <- metrics_fit1_test[metrics_fit1_test$.metric == "rmse", ".estimate"]

@@

When running the model on the testing dataset, the results remain significant at the 0.05 level. The RMSE is `r round(rmse_1_test$model1_test)` and the r-squared is `r round(r2_1_test$model1_test, 3)`. This shows a slightly worse fit than the training results.

@@

whereas `BW` is birth weight in grams, $\alpha$ is the average birth weight, $\beta_x$ are the group deviations from the average birth weight, `MA` is maternal ancestry (Euro-descent, Afro-descent), `Gest` is gestational status (nulliparous or multiparous), `Age` is maternal age in years, and $\epsilon$ is the error term. Results are displayed in Table 3.

```{r, message=FALSE,  warning=FALSE}
# Model 2: Final model for paper using all data
# Create a list to store R-squared values
r2_2 <- list()
rmse_2 <- list()

# Using recipes and workflows
# Define recipe for model
ML_recipe2 <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") #%>% # Set reference level for ModifiedColor
  #step_dummy(ModifiedColor) #%>% # Convert all nominal variables to dummy variables. Cannot do this if I am merging tables in gtsummary
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors. Cannot do this if I am merging tables in gtsummary

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit2_pred <- bind_cols(ML_fit2_pred, ML_linear %>% select(Weightgrams))
ML_fit2_pred

@@

# Extract RMSE and R2 for in-line text
r2_2$model2 <- metrics_fit2[metrics_fit2$.metric == "rsq", ".estimate"]
rmse_2$model2 <- metrics_fit2[metrics_fit2$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model2, file = here::here("results", "tables", "table_model2_final.rds"))

# Create a gtsummary table directly from the model with CIs
table_model2_CI <- tbl_regression(ML_fit2, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model2_CI
```

The RMSE is `r round(rmse_2$model2)` and the r-squared is `r round(r2_2$model2, 3)`.

@@

Now, I'll run the model on the training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 2
# Create a list to store R-squared and RMSE values
r2_2_train <- list()
rmse_2_train <- list()

# Using recipes and workflows
# Define recipe for model
ML_recipe2_train <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit2_pred_train <- bind_cols(ML_fit2_pred_train, train_data %>% select(Weightgrams))
ML_fit2_pred_train

@@

# Extract RMSE and R2 for in-line text
r2_2_train$model2_train <- metrics_fit2_train[metrics_fit2_train$.metric == "rsq", ".estimate"]
rmse_2_train$model2_train <- metrics_fit2_train[metrics_fit2_train$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model2_train, file = here::here("results", "tables", "table_model2_train.rds"))
```
The RMSE is `r round(rmse_2_train$model2_train)` and the r-squared is `r round(r2_2_train$model2_train, 3)` a slightly better performance than the full model.

Now I'll run cross-validation.

@@

# Create a list to store R-squared and RMSE values
r2_2_cv <- list()
rmse_2_cv <- list()

@@

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_2_cv$model2_cv <- ML_cv2_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_2_cv$model2_cv <- ML_cv2_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv2_metrics, file = here::here("results", "tables", "ML_cv2_metrics.rds"))
```
Here, the RMSE is `r round(rmse_2_cv$model2_cv)` and the r-squared is `r round(r2_2_cv$model2_cv, 3)`, the same as the training model `r round(rmse_2_train$model2_train)`, `r round(r2_2_train$model2_train, 3)`.

Finally, I'll run the model on the test data.

```{r, message=FALSE,  warning=FALSE}
# Testing model 2
# Create a list to store R-squared and RMSE values
r2_2_test <- list()
rmse_2_test <- list()

# Using recipes and workflows
# Define recipe for model
ML_recipe2_test <- recipe(Weightgrams ~ ModifiedColor + Age + ModifiedStatus + ModifiedNationality, data = test_data) %>% 
  step_relevel(ModifiedColor, ref_level = "Euro-Descent") %>% # Set reference level for ModifiedColor
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit2_pred_test <- bind_cols(ML_fit2_pred_test, test_data %>% select(Weightgrams))
ML_fit2_pred_test

@@

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_2_test$model2_test<- metrics_fit2_test[metrics_fit2_test$.metric == "rsq", ".estimate"]
rmse_2_test$model2_test <- metrics_fit2_test[metrics_fit2_test$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model2_test, file = here::here("results", "tables", "table_model2_test.rds"))
```

Running the model on the test data, there are no significant results at any alpha level.

@@

```{r, message=FALSE,  warning=FALSE}
# Model 1: Sub-sample simple model (ML_lm1_sub)
# Create a list to store R-squared values
r_squared_values_sub <- list()

@@

# Save regression output into table
saveRDS(ML_lm1_sub_table, file = here::here("results", "tables", "ML_lm1_sub_table.rds"))
```
```{r, message=FALSE,  warning=FALSE}
# Model 2: Sub-sample multivariate model (ML_lm2_sub)
# Filter data based on desired conditions
filtered_data <- ML_linear %>%
  filter(Brazilian == "Brazilian")

@@

```{r, message=FALSE,  warning=FALSE}
# Combine the four tables using tbl_merge
table3 <- tbl_merge(
  list(table_model1, ML_lm1_sub_table, table_model2, ML_lm2_sub_table),
  tab_spanner = c("**Model 1**", "**Sub-Model 1**", "**Model 2**", "**Sub-Model 2**")) %>%
  modify_table_body(fun = ~.x %>% arrange(variable))
table3

@@

```{r, message=FALSE,  warning=FALSE}
# Model 3: Final model for paper using all data
# Create a list to store R-squared and RMSE values
r2_3 <- list()
rmse_3 <- list()

# Using recipes and workflows
# Simple linear model
# Specify recipe
ML_recipe3 <- recipe(Weightgrams ~ Color, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") #%>% # Set reference level for White
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit3_pred <- bind_cols(ML_fit3_pred, ML_linear %>% select(Weightgrams))
ML_fit3_pred

@@

# Extract RMSE and R2 for in-line text
r2_3$model3 <- metrics_fit3[metrics_fit3$.metric == "rsq", ".estimate"]
rmse_3$model3 <- metrics_fit3[metrics_fit3$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model3, file = here("results", "tables", "table_model3_final.rds"))

# Create a gtsummary table from model with CIs
table_model3_CI <- tbl_regression(ML_fit3, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  add_glance_table(include = c(r.squared, nobs)) %>%
  modify_header(statistic = "**T-Statistic**") # rename header from statistic to t-statistic
table_model3_CI
```

The RMSE is `r round(rmse_3$model3)` and the r-squared is `r round(r2_3$model3, 3)`.

@@

Now I'll evaluate the model on our training data.

```{r, message=FALSE,  warning=FALSE}
# Training model 3
# Create a list to store R-squared and RMSE values
r2_3_train <- list()
rmse_3_train <- list()

# Using recipes and workflows
# Specify recipe
ML_recipe3_train <- recipe(Weightgrams ~ Color, data = train_data) %>%
  step_relevel(Color, ref_level = "White")

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit3_pred_train <- bind_cols(ML_fit3_pred_train, train_data %>% select(Weightgrams))
ML_fit3_pred_train

@@

# Extract RMSE and R2 for in-line text
r2_3_train$model3_train <- metrics_fit3_train[metrics_fit3_train$.metric == "rsq", ".estimate"]
rmse_3_train$model3_train <- metrics_fit3_train[metrics_fit3_train$.metric == "rmse", ".estimate"]

@@

The RMSE is `r round(rmse_3_train$model3_train)` and the r-squared is `r round(r2_3_train$model3_train, 3)`, slightly better performance than the full model `r round(r2_3$model3, 3)`. The results are not significantly different from the full model.

Now I'll run cross-validation.

@@

# Create list to store R-squared and RMSE values
r2_3_cv <- list()
rmse_3_cv <- list()

@@

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_3_cv$model3_cv <- ML_cv3_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_3_cv$model3_cv <- ML_cv3_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv3_metrics, file = here::here("results", "tables", "ML_cv3_metrics.rds"))
```
Using cross-validation, the RMSE is `r round(rmse_3_cv$model3_cv)` and the r-squared is `r round(r2_3_cv$model3_cv, 3)`, similar performances to the training model `r round(rmse_3_train$model3_train)`, `r round(r2_3_train$model3_train, 3)`.

Finally, I'll run the model on the test data.

```{r, message=FALSE, warning=FALSE}
# Testing model 3
# Create a list to store R-squared and RMSE values
r2_3_test <- list()
rmse_3_test <- list()

# Using recipes and workflows
# Specify recipe
ML_recipe3_test <- recipe(Weightgrams ~ Color, data = test_data) %>%
  step_relevel(Color, ref_level = "White")

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit3_pred_test <- bind_cols(ML_fit3_pred_test, test_data %>% select(Weightgrams))
ML_fit3_pred_test

@@

# Extract R-squared and RMSE values for in-line text (because data frame with columns not list)
r2_3_test$model3_test<- metrics_fit3_test[metrics_fit3_test$.metric == "rsq", ".estimate"]
rmse_3_test$model3_test <- metrics_fit3_test[metrics_fit3_test$.metric == "rmse", ".estimate"]

@@

Much like in the other testing models, the results are not significant (except the intercept). The performance is also worse (RMSE of `r round(rmse_3_test$model3_test)` and r-squared of `r round(r2_3_test$model3_test, 3)`).

@@

```{r, message=FALSE,  warning=FALSE}
# Model 4: Final model for paper using all data
# Create a list to store R-squared values
r2_4 <- list()
rmse_4 <- list()

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4 <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = ML_linear) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") #%>% # Set reference level for ModifiedStatus
  #step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  #step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit4_pred <- bind_cols(ML_fit4_pred, ML_linear %>% select(Weightgrams))
ML_fit4_pred

@@

# Extract RMSE and R2 for in-line text
r2_4$model4 <- metrics_fit4[metrics_fit4$.metric == "rsq", ".estimate"]
rmse_4$model4 <- metrics_fit4[metrics_fit4$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model4, file = here("results", "tables", "table_model4_final.rds"))

# Create a gtsummary table from model with CIs
table_model4_CI <- tbl_regression(ML_fit4, exponentiate = FALSE, intercept = TRUE) %>%
  modify_column_unhide(columns = c(statistic, std.error)) %>%
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>% # Hide confidence intervals and p-values
  add_glance_table(include = c(r.squared, nobs)) %>% # Add R-squared and number of observations
  modify_header(statistic = "**T-Statistic**")
table_model4_CI
```
The RMSE is `r round(rmse_4$model4)` and the r-squared is `r round(r2_4$model4, 3)`.

@@

I will now evaluate the model on the training data.

```{r, message=FALSE,  warning=FALSE}
# Model 4 training
# Create list to store R-squared and RMSE values
r2_4_train <- list()
rmse_4_train <- list()

# Using recipes and workflows
# Complex linear model
# Specify recipe
ML_recipe4_train <- recipe(Weightgrams ~ Color + Age + ModifiedStatus + ModifiedNationality, data = train_data) %>%
  step_relevel(Color, ref_level = "White") %>% # Set reference level for Color
  step_relevel(ModifiedStatus, ref_level = "Multiparous") %>% # Set reference level for ModifiedStatus
  step_dummy(all_nominal(), -all_outcomes()) %>% # Convert all nominal variables to dummy variables
  step_normalize(all_numeric_predictors()) # Normalize all numeric predictors

@@

# Predicted numeric outcome named .pred. Match predicted values with corresponding observed outcome values
ML_fit4_pred_train <- bind_cols(ML_fit4_pred_train, train_data %>% select(Weightgrams))
ML_fit4_pred_train

@@

# Extract RMSE and R2 for in-line text
r2_4_train$model4_train <- metrics_fit4_train[metrics_fit4_train$.metric == "rsq", ".estimate"]
rmse_4_train$model4_train <- metrics_fit4_train[metrics_fit4_train$.metric == "rmse", ".estimate"]

# Save regression output into table
saveRDS(table_model4_train, file = here("results", "tables", "table_model4_train.rds"))
```

The RMSE is `r round(rmse_4_train$model4_train)` and the r-squared is `r round(r2_4_train$model4_train, 3)` a slightly better performance than the full model `r round(r2_4$model4, 3)`.

Now I'll run cross-validation.

@@

# Create list to store R-squared and RMSE values
r2_4_cv <- list()
rmse_4_cv <- list()

@@

# Extract R-squared and RMSE values for in-line text
# For R-squared
r2_4_cv$model4_cv <- ML_cv4_metrics %>%
  filter(.metric == "rsq") %>%
  pull(mean)

# For RMSE
rmse_4_cv$model4_cv <- ML_cv4_metrics %>%
  filter(.metric == "rmse") %>%
  pull(mean)

# Save regression output into table
saveRDS(ML_cv4_metrics, file = here::here("results", "tables", "ML_cv4_metrics.rds"))
``` 
Here, the RMSE is `r round(rmse_4_cv$model4_cv)` and the r-squared is `r round(r2_4_cv$model4_cv, 3)`, a better performance than both the full model `r round(rmse_4$model4)`, `r round(r2_4$model4, 3)` and the training model `r round(r2_4_train$model4_train, 3)`, `r round(rmse_4_train$model4_train)`.

Finally, I will run model on the test data.
